---
title: "pmlProject"
author: "abe hashemi"
date: "Sunday, July 26, 2015"
output: html_document
---

 
Load necessary libraries:


```{r}
library(caret)
library(randomForest)
library(rpart)     # For regression trees and partition
library(rpart.plot) 	# For plotting Decision Tree 

```



 set the overall seed for reproduceability
 
 

```{r}
set.seed(1234)


```



Create project data directory, if none exists:

```{r}
if(!file.exists("./pData")){dir.create("./pData")}

```

set the data file urls:

```{r}
dTrainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
dTestUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

```

load the data files:

```{r}
pmlTrainDs <- read.csv(url(dTrainUrl), na.strings=c("NA","#DIV/0!", ""))
pmlTestDs <- read.csv(url(dTestUrl))


```

 
Check the structure of the data sets:

```{r}
str(pmlTrainDs)
str(pmlTestDs)


```



Clean out all NA variables:
```{r}

pmlTrainDs <- pmlTrainDs[,colSums(is.na(pmlTrainDs)) == 0]
pmlTestDs <- pmlTestDs[,colSums(is.na(pmlTestDs)) == 0]

```

Clean out columns 1-7 as irrelevant to this analysis:
```{r}
pmlTrainDs <- pmlTrainDs[,-c(1:7)]
pmlTestDs <- pmlTestDs[,-c(1:7)]

```

The clean data:

```{r}
str(pmlTrainDs)
str(pmlTestDs)

```

Divide training data into a subTrain and SubTest random samples of 70:30 size ratios:
```{r}
subtrainChoice <- createDataPartition(y=pmlTrainDs$classe, p=0.70, list=FALSE)
pmlSubtrainDs <- pmlTrainDs[subtrainChoice,]
pmlSubtestDs <- pmlTrainDs[-subtrainChoice,]

```
The subtrain and subtest data:
```{r}

str(pmlTrainDs)
str(pmlTestDs)

```

Model first based on Decision Tree Classification:

```{r}
dcPmlModel <- rpart(classe ~ ., data=pmlSubtrainDs, method="class")
rpart.plot(dcPmlModel, main="PML Classification Tree", extra=107)  # Plot the tree

```

Predict the subtest data set based on this decision tree:

```{r}
dcPmlPrediction <- predict(dcPmlModel, pmlSubtestDs, type = "class")

```

Get the confusion matrix for decision tree:

```{r}
dcPmlCm <- confusionMatrix(dcPmlPrediction, pmlSubtestDs$classe)

```

# Now let's look at decision forests for more accuracy:
```{r}
dfPmlModel <- randomForest(classe ~. , data=pmlSubtrainDs, method="class")

```

Predict the subtest data set based on this decision forest:

```{r}
dfPmlPredition <- predict(dfPmlModel, pmlSubtestDs, type = "class")

```

Get the confusion matrix for decision tree model:

```{r}
confusionMatrix(dfPmlPredition, pmlSubtestDs$classe)

```
#Conclusion
Apply the better accuracy decision forest to the test data set:
predict outcome levels on the original Testing data set using Random Forest algorithm

```{r}
pmlPredict <- predict(dfPmlModel, pmlTestDs, type="class")

```


# Save results:
Write files for submission:

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pmlPredict)

```

# References:
[1] Velloso, E.; Bulling, A.; Gellersen, H.; Ugulino, W.; Fuks, H. Qualitative Activity Recognition of Weight Lifting Exercises. Proceedings of 4th International Conference in Cooperation with SIGCHI (Augmented Human '13) . Stuttgart, Germany: ACM SIGCHI, 2013.

[2] Krzysztof Gra??bczewski and Norbert Jankowski. Feature Selection with Decision Tree Criterion.